Do larger watersheds respond different to forest cover change? Re-analysing a global data set.

### 2021-12-02 paper/Forest_and_Water/Forest_and_Water.tex
updates paper 03122021

In addition, additional variables added were the latitude and longitude
for the center of the watershed as an approximation of its spatial
location. Using this information reference evapotranspiration (E\_\{0\})
was extracted from the Global Aridity Index and Potential
Evapo-Transpiration (ET0) Climate Databasev2 (Trabucco and Zomer, 2018),
if a value of E0 was not available from the original papers. For large
watersheds, this value, similar to annual average rainfall, is only an
approximation of the climate at the location.

@@

Similar to Zhang et al. (2017), the ``dryness index'' was calculated as:
\[\tag{1}
\begin{aligned}
D = \frac{E0}{Pa}
\end{aligned}\]

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Zhang_all <-}\StringTok{ }\NormalTok{Zhang_all }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\StringTok{`}\DataTypeTok{Watershed #}\StringTok{`} \OperatorTok{>}\StringTok{ }\DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Zhang_all <-}\StringTok{ }\NormalTok{Zhang_all }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{DeltaQf_perc =} \KeywordTok{ifelse}\NormalTok{(}\StringTok{`}\DataTypeTok{Watershed #}\StringTok{`} \OperatorTok{==}\StringTok{ }\DecValTok{76}\NormalTok{,}\DecValTok{157}\NormalTok{,DeltaQf_perc))}
\end{Highlighting}
\end{Shaded}

@@

Here \(X_i\) are factorial variables, while \(Z_i\) are continous
variables. The model assumes no direct interactions and all variables
are additive. The changes in forest cover contain both positive
(forestation) and negative values (deforestation). In Zhang et al.
(2017), these changes were jointly analysed, assuming the effect on the
change in flow was linear and non-hysteretic. However, the impact of an
increase in forest cover can be different from the same fractional
decrease in forest cover. Therefore all the change in forest cover data
is converted to positive values, and an additional column
(\emph{sign\_\{forest cover\}}) is added that indicates whether it was a
forest cover increase or decrease. A further assumption in the model is
that all continuous variables \(Z_i\) (such as annual precipitation
(Pa)) can have a linear or non-linear relationship with \(\Delta Q \%\).
This means that a smooth function \(s()\) is applied to the \(Z_i\)
variables.

@@

The overall dataset contains 311 observations of changes in flow. The
overall distribution of changes in flow is highly skewed as is the
distribution of changes in forest cover and Area. The values of changes
in flow greater than 100\% and smaller than -100\% clearly create long
tails on the change in flow distribution. Note also the large number of
studies with 100\% forest cover reduction. Smaller watersheds dominate
the database with 42\% of the data from watersheds \textless{} 1
km\textsuperscript{2} and 65\% of the data for watersheds \textless{} 10
km\textsuperscript{2}.

@@

\begin{longtable}[]{@{}ccccc@{}}
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{(Intercept)}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
8.77\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
5.52\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
1.59\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.11\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{DeltaF\_perc\_pos}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.5\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0.09\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
5.77\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Forest\_Signincrease}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-30.9\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
5.86\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-5.27\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

While the overall variance explained in this model is not high with an
adjusted \emph{r\textsuperscript{2}} of 0.19, it clearly supports the
hypothesized relationship between the change in forest cover and the
change in flow. The model suggests that for every 1\% change in forest
cover, on the average, the flow changes 0.5\%. However the change in
flow is different for forest cover decreases compared to forest cover
increases. In fact, forest cover increases decrease flow by 31\% less
than a similar decrease in forest cover causes flow to increase. So
roughly speaking, a 1\% forest cover increase on the average decreases
flow by \((1 - 0.31)*0.5\%\), while a the percentage forest cover
decrease will increase flow by 0.5\%.

It is however it is clear from the lack of explaining power, that there
could be confounding factors, as alluded to in the methods. The obvious
ones being watershed dryness and area (following Zhang et al. (2017)):

\[\tag{4}
\begin{aligned}
\Delta \% Q \sim ~ &\Delta \% forest~cover_{positive} + sign_{forest~cover} + \\ & Pa_{mm} + Area_{km^2} + \varepsilon
\end{aligned}\] Where \(Pa_mm\) is the annual average rainfall in mm.

\begin{longtable}[]{@{}ccccc@{}}
\caption{Summary of the second model, taking into account the annual
rainfall and the area of the watershed}\tabularnewline
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{(Intercept)}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
18.94\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
7.78\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
2.43\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.02\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{DeltaF\_perc\_pos}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.5\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0.09\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
5.66\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Forest\_Signincrease}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-31.54\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
5.9\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-5.35\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Area\_km2}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-0.3\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.77\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Pa\_mm}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-0.01\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-1.75\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.08\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

Including area and annual precipitation does not really improve the
overall explaining power of the model, in fact, annual precipitation
appears to be only a very small confounding factor, representing only a
-0.01/\% partial effect in the change in streamflow, holding all other
factors constant. In contrast to earlier reported studies (Filoso et
al., 2017; Zhang et al., 2017), watershed area has no effect on the
change in stream flow. This supports our approach (in contrast to Zhang
et al. (2017)) to consider watershed area as a continuous variable and
making no separation between larger and smaller watersheds The main
effects remain the change in forest cover and whether this is an
increase or decrease.

@@

\[\tag{5}
\begin{aligned}
\Delta \% Q \sim ~ &\Delta \% forest~cover_{positive} + sign_{forest~cover} + \\ & Pa_{mm} + Area_{km^2} + Latitude + Longitude + \varepsilon
\end{aligned}\]

\begin{longtable}[]{@{}ccccc@{}}
\caption{Results of the model including Latitude and
Longitude}\tabularnewline
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{(Intercept)}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
29.99\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
8.65\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
3.47\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{DeltaF\_perc\_pos}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.47\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0.09\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
5.4\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Forest\_Signincrease}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-37.11\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
6.09\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-6.09\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Area\_km2}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-0.59\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.55\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Pa\_mm}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-0.01\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-2.17\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.03\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Latitude}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-0.29\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0.11\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-2.74\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.01\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Longitude}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.01\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0.03\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
0.28\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.78\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

>>>>

In addition, additional variables added were the latitude and longitude
for the center of the watershed as an approximation of its spatial
location. Using this information reference evapotranspiration
(E\textsubscript{0}) was extracted from the Global Aridity Index and
Potential Evapo-Transpiration (ET\textsubscript{0}) Climate Databasev2
(Trabucco and Zomer, 2018), if a value of E\textsubscript{0} was not
available from the original papers. For large watersheds, this value,
similar to annual average rainfall, is only an approximation of the
climate at the location.

@@

Similar to Zhang et al. (2017), the ``dryness index'' was calculated
from the reference evapotranspiration and the annual average rainfall
(Pa) as: \[\tag{1}
\begin{aligned}
D = \frac{E_{0}}{Pa}
\end{aligned}\]

@@

Removing possible duplicates removes 29 watersheds from the overall data
set

@@

Here \(X_i\) are factorial variables, while \(Z_i\) are continous
variables. The model assumes no direct interactions and all variables
are additive. The changes in forest cover contain both positive
(forestation) and negative values (deforestation). In Zhang et al.
(2017), these changes were jointly analysed, assuming the effect on the
change in flow was linear and non-hysteretic. However, the impact of an
increase in forest cover can be different from the same fractional
decrease in forest cover. Therefore all the change in forest cover data
is converted to positive values, and an additional column
(\emph{sign\_\{forest cover\}}) is added that indicates whether it was a
forest cover increase or decrease. A further assumption in the model is
that all continuous variables \(Z_i\) (such as annual precipitation
(Pa)) can have a linear or non-linear relationship with \(\Delta Q \%\).
This means that a smooth function \(s()\) is applied to the \(Z_i\)
variables. For the smoothing function we applied thin plate regression
splines with an additional shrinkage penalty which means the terms can
be shrunk to 0 if not significant (Wood, 2006).

@@

The overall dataset contains 312 observations of changes in flow. The
overall distribution of changes in flow is highly skewed as is the
distribution of changes in forest cover and Area. The values of changes
in flow greater than 100\% and smaller than -100\% clearly create long
tails on the change in flow distribution. Note also the large number of
studies with 100\% forest cover reduction. Smaller watersheds dominate
the database with 42\% of the data from watersheds \textless{} 1
km\textsuperscript{2} and 65\% of the data for watersheds \textless{} 10
km\textsuperscript{2}.

@@

\begin{longtable}[]{@{}ccccc@{}}
\caption{Summary results of the first regression model predicting change
in streamflow from change in forest cover and accounting for the
direction of the change}\tabularnewline
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{(Intercept)}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
8.65\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
5.56\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
1.56\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.12\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{DeltaF\_perc\_pos}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.45\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0.09\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
5.26\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Forest\_Signincrease}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-29.17\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
5.79\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-5.04\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}ccccc@{}}
\caption{Summary results of the first regression model predicting change
in streamflow from change in forest cover and accounting for the
direction of the change including the new data sets}\tabularnewline
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{(Intercept)}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
9.43\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
5.54\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
1.7\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.09\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{DeltaF\_perc\_pos}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.44\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0.09\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
5.12\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Forest\_Signincrease}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-36.54\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
5.59\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-6.53\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

While the overall variance explained in this model is not high with an
adjusted \emph{r\textsuperscript{2}} of 0.18, it clearly supports the
hypothesized relationship between the change in forest cover and the
change in flow. The model suggests that for every 1\% change in forest
cover, on the average, the flow changes 0.45\%. However the change in
flow is different for forest cover decreases compared to forest cover
increases. In fact, forest cover increases decrease flow by 29\% less
than a similar decrease in forest cover causes flow to increase. So
roughly speaking, a 1\% forest cover increase on the average decreases
flow by \((1 - 0.29)*0.45\%\), while a the percentage forest cover
decrease will increase flow by 0.45\%.

Checking this result against including the new data indicates that this
mainly strengthens the difference between the forest cover increases and
decreases, and decreases the mean decrease in flow as a result of forest
cover change

It is however it is clear from the lack of explaining power for the
model, that there could be confounding factors, as alluded to in the
methods. The obvious ones being watershed dryness and area (following
Zhang et al. (2017)):

\[\tag{4}
\begin{aligned}
\Delta \% Q \sim ~ &\Delta \% forest~cover_{positive} + sign_{forest~cover} + \\ & Pa_{mm} + log10(Area_{km^2}) + \varepsilon
\end{aligned}\]

Where \(Pa_mm\) is the annual average rainfall in mm, and a log base 10
transformation is applied to the area variable given that the
distribution of Area is higly skewed (Figure 1)

\begin{longtable}[]{@{}ccccc@{}}
\caption{Summary of the second model, taking into account the annual
rainfall and the area of the watershed}\tabularnewline
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{(Intercept)}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
22.55\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
9.16\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
2.46\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.01\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{DeltaF\_perc\_pos}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.34\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0.1\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
3.26\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Forest\_Signincrease}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-35.52\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
5.67\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-6.27\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{log10(Area\_km2)}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-3.12\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
1.72\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-1.81\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.07\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Pa\_mm}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-1\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.32\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

Including area and annual precipitation slightly improves the overall
explaining power of the model. Annual precipitation is in fact not
significant. Relative to earlier reported studies (Filoso et al., 2017;
Zhang et al., 2017), the log base 10 transformed watershed area
indicates a p-value of only 0.07, suggesting a marginal impact on the
change in stream flow. This supports our approach (in contrast to Zhang
et al. (2017)) to consider watershed area as a continuous variable and
making no separation between larger and smaller watersheds The main
effects remain the change in forest cover and whether this is an
increase or decrease.

@@

\[\tag{5}
\begin{aligned}
\Delta \% Q \sim ~ &\Delta \% forest~cover_{positive} + sign_{forest~cover} + \\ & Pa_{mm} + log10(Area_{km^2}) + Latitude + Longitude + \varepsilon
\end{aligned}\]

\begin{longtable}[]{@{}ccccc@{}}
\caption{Results of the model including Latitude and Longitude including
new data}\tabularnewline
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.31\columnwidth}\centering
~\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Estimate\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\centering
Std. Error\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\centering
t value\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
Pr(\textgreater\textbar t\textbar)\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{(Intercept)}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
23.29\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
9.96\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
2.34\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.02\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{DeltaF\_perc\_pos}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.35\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0.1\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
3.32\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Forest\_Signincrease}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-37.21\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
6\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-6.2\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{log10(Area\_km2)}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-3.28\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
1.73\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-1.89\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.06\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Pa\_mm}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-0.83\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.41\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Latitude}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
-0.05\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0.09\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
-0.55\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.58\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\centering
\textbf{Longitude}\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.01\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\centering
0.03\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\centering
0.23\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
0.82\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

There appears to be no significant latitudinal gradient. The total
explaining power of the model is still low with an adjusted
\emph{r\textsuperscript{2}} of 0.22 suggesting further factors that are
currently not included in the model.
